%
\documentclass[phd,tocprelim]{cornell}

% tocprelim option must be included to put the roman numeral pages in the
% table of contents
%
% The cornellheadings option will make headings completely consistent with
% guidelines.
%
% This sample document was originally provided by Blake Jacquot, and
% fixed up by Andrew Myers.
%
%Some possible packages to include
\usepackage{graphicx,pstricks}
\usepackage{graphics}
\usepackage{moreverb}
%\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{hangcaption}
\usepackage{palatino}
\usepackage{textcomp}
\usepackage{gensymb}
\usepackage{natbib}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{csvsimple}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{cleveref}

\usepackage{subcaption}
\captionsetup{format=plain}

\def\code#1{\texttt{#1}}
%if you're having problems with overfull boxes, you may need to increase
%the tolerance to 9999
\tolerance=9999

\bibliographystyle{abbrvnat}
% \setcitestyle{authoryear,open={((},close={))}} %Citation-related commands
% \bibliographystyle{plain}
% \bibliographystyle{natbib}

\renewcommand{\caption}[1]{\singlespacing\hangcaption{#1}\normalspacing}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\title {Planning direct imaging observations of exoplanets with precursor data}
\author {Corey Spohn}
\conferraldate {August}{2023}
\degreefield {Ph.D.}
\copyrightholder{Corey Spohn}
\copyrightyear{2023}

\begin{document}

\maketitle
\makecopyright

\begin{abstract}
Exoplanets have become a major focus in astronomy over the last two decades which has
driven a rapid increase in the number of detected exoplanets. The detection of new
exoplanets will drive the future of astronomy. Our understanding of planetary
formation, the conditions for extraterrestrial life, and the location of
habitable worlds are directly dependent on obtaining high-quality observations
of exoplanets. There are a number of methods used to detect an exoplanet
through direct and indirect observations. Direct detection of an exoplanet is
done by observing the exoplanet itself, while indirect detection is done by
observing changes in the star or circumstellar disk. No single detection method can determine
everything about an exoplanet, and each is biased toward certain kinds of
planets. The work presented here focuses on the problem of directly
detecting exoplanets that have been discovered indirectly 
via the ``radial velocity'' method.

The first part of the dissertation focuses on how to calculate the times when a
planet with an orbit fitted via the radial velocity detection method will be
detectable for a direct imaging telescope. The dissertation analyzes how to convert the orbit
fits, estimate the probability of detection as a function of time, and identify
two major failure modes inherent to the process.

The second part of the dissertation identifies a problem with a traditionally
used method of calculating the dimmest planet that can be detected for a given
integration time, the observing time spent collecting photons for science. In
creating higher fidelity models of the signal to noise ratio for exoplanet
detection, the signal and noise terms have become inseparable which makes it
impossible to calculate the dimmest planet detectable as a function of
integration time analytically. The dissertation demonstrate a numerical routine
to solve this problem and use it to calculate the performance of the Roman
Space Telescope's different observing scenarios on known exoplanets.

Finally, the dissertation uses the numerical routine to determine the dimmest
planet detectable to create a more accurate model to estimate the probability
of directly imaging an exoplanet detected via the radial velocity method by
accounting for additional sources of noise. The probability is found by using
the dimmest planet detectable over a direct imaging instrument's possible
planet-star separations, local zodiacal light values, and extra zodiacal light
values and comparing those values to the estimated positions and brightnesses
of planets consistent with the underlying radial velocity data. The probability
of detection metric is validated by creating a framework for full simulations
of the process. The framework generates a set of radial velocity data for
synthetic exoplanets, fits orbits to the radial velocity data, calculates the
probability of detection, schedules observations, and simulates the
observations. The scheduling algorithm is validated in a simulation of future
radial velocity and direct imaging instruments. In the simulation the
scheduling algorithm is compared against random searches and the algorithm, on
average, detects $\sim$32 unique planets in $\sim$375 days of integration time
while a random walk detects $\sim$17 unique planets in $\sim$720 days of
integration time. Finding twice as many planets in half the time is a
significant improvement and proves that direct imaging observations of
previously detected planets can be reliably scheduled with the assumed
instrument performance and mission design parameters.


\end{abstract}

\begin{biosketch}
Corey Spohn grew up in Reston, Virginia and graduated from Herndon High School
in 2013. He attended Virginia Tech, graduating in 2018 with a BS in physics and
a second BS in engineering science and mechanics because he couldn't make up
his mind. At Virginia Tech he was exposed to engineering research when he received
an email looking for people to help with flying snake research. Working with
flying snakes taught him to never do research on live animals, so he then
joined the Nature Inspired Fluids and Interfaces lab where he spent hours
watching water droplets freeze which was oddly fulfilling. It instilled in
Corey an interest in research and convinced him to apply to graduate school. In
2018 he began as a Cornell PhD student and joined the Space Imaging and Optical
Systems lab, working with Professor Dmitry Savransky, to work on exoplanet
mission planning.
\end{biosketch}

\begin{dedication}
To my family of the past, present, and future.
\end{dedication}

\begin{acknowledgements}

I am incredibly fortunate to have had so many wonderful people to led me to
this point in my life. I don't know how to thank my parents enough for all the
love and support they have given to me, I would not be who I am without you.
Thanks to my family, particularly my grandmother and my siblings. Thank you
Abigail Brown for making Ithaca home. Thanks to all the teachers I have had,
the two that stand out are Mr. Ence for inspiring me to pursue physics and
Jonathan Boreyko for all the effort he gave to mentor undergrad researchers.

I am forever indebted to Dmitry Savransky for accepting me into his lab and for
the countless hours he has spent helping me. I have become a better researcher,
programmer, presenter, and writer thanks to your guidance. I aspire to be a
researcher with the breadth of knowledge you have, and I cannot overstate how
high I believe that bar is. Thanks to my labmates Gabe, Jacob, Dean, Duan,
Grace, Rachel, Jackson, Zvonimir, and Colby for all the help and fun times in
the lab. Special thanks to Dean and Grace for all the messages we've exchanged
about EXOSIMS. Cornell MAE's graduate student community made my time in Ithaca
rewarding and memorable, and the community would not have been the same without
Marcia Sawyer. Thanks to all my collaborators, particularly Rhonda Morgan.

Finally, I'd like to say thanks to all the close friends I've made in Ithaca
and Blacksburg, in particular thanks to Jimmy, Rob, James, David, Kalani, and
Grant for the good times we've had.

% This work was funded by the Science Investigation Team of the Nancy Grace Roman
% Space Telescope under NASA Grant NNX15AB40G.

\end{acknowledgements}

\contentspage
\tablelistpage
\figurelistpage

\normalspacing \setcounter{page}{1} \pagenumbering{arabic}
\pagestyle{cornell} \addtolength{\parskip}{0.5\baselineskip}

\chapter{Introduction}
\label{cha:intro}
\input{intro/intro.tex}

\chapter{Scheduling Direct Imaging Observations Based on Radial Velocity Orbital Fits: Best Practices for Translating Orbits and Failure Modes}
\label{cha:first_paper}
\input{ch1/ch1.tex}

\chapter{Coupling of Planet and Background Signal}
\label{cha:coupling}
\input{ch2/coupling.tex}

\chapter{Accurate Estimation of Probability of Imaging Detection with Precursor Radial Velocity Data}
\label{cha:accurate_pdet}
\input{ch3/ch3.tex}

\chapter{Scheduling with Simulated Precursor Data Collection for Direct Imaging Yield Modeling}
\label{cha:sim_and_scheduling}
\input{ch4/ch4.tex}

\chapter{Conclusion}
\label{cha:conclusion}
\input{conclusion/conclusion.tex}


\appendix
% \chapter{Chapter 1 of appendix}
% Appendix chapter 1 text goes here

\bibliography{./library.bib}

\end{document}
